---
title: "Plotly"
author: "Edward Coleman"
date: "West Chester University"
output:
  html_document: 
    code_folding: hide
    code_download: yes
    smooth_scroll: yes
    theme: lumen
editor_options:
  chunk_output_type: inline
---

<style type="text/css">

div#TOC li {
    list-style:none;
    background-color:lightgray;
    background-image:none;
    background-repeat:none;
    background-position:0;
    font-family: Arial, Helvetica, sans-serif;
    color: #780c0c;
}

h1.title {
  font-size: 24px;
  color: DarkRed;
  text-align: center;
  font-family: Arial, Helvetica, sans-serif;
  font-variant-caps: normal;
}
h4.author { 
    font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkRed;
  text-align: center;
}
h4.date { 
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
  text-align: center;
}
h1 { 
    font-size: 22px;
    font-family: "Times New Roman", Times, serif;
    color: darkred;
    font-weight: bold;
    text-align: center;
}
h2 { 
    font-size: 18px;
    font-family: "Times New Roman", Times, serif;
    color: navy;
    text-align: left;
}

h3 { 
    font-size: 18px;
    font-family: "Times New Roman", Times, serif;
    color: navy;
    font-weight: bold;
    text-align: left;
}

h4 {
    font-size: 18px;
    font-family: "Times New Roman", Times, serif;
    color: darkred;
    text-align: left;
}


/* Tab features */
.nav>li>a {
    position: relative;
    display: block;
    padding: 2px 15px;
    color: #990000;
}
.nav-pills>li.active>a, .nav-pills>li.active>a:hover, .nav-pills>li.active>a:focus {
    color: #ffffff;
    background-color: #990000;
}
/*
nav-pills>li:nth-child(2) {
    background: green;
 }
 */
</style>

```{r setup, include=FALSE, comment=NA}
options(repos = list(CRAN="http://cran.rstudio.com/"))
if (!require("tidyverse")) {
   install.packages("tidyverse")
   library(tidyverse)
}
if (!require("knitr")) {
   install.packages("knitr")
   library(knitr)
}
if (!require("cowplot")) {
   install.packages("cowplot")
   library(cowplot)
}
if (!require("latex2exp")) {
   install.packages("latex2exp")
   library(latex2exp)
}
if (!require("plotly")) {
   install.packages("plotly")
   library(plotly)
}
if (!require("gapminder")) {
   install.packages("gapminder")
   library(gapminder)
}
if (!require("png")) {
    install.packages("png")    
    library("png")
}
if (!require("RCurl")) {
    install.packages("RCurl")    
    library("RCurl")
}
if (!require("colourpicker")) {
    install.packages("colourpicker")              
    library("colourpicker")
}
if (!require("gganimate")) {
    install.packages("gganimate")              
    library("gganimate")
}
if (!require("gifski")) {
    install.packages("gifski")              
    library("gifski")
}
if (!require("magick")) {
    install.packages("magick")              
    library("magick")
}
if (!require("grDevices")) {
    install.packages("grDevices")              
    library("grDevices")
}
if (!require("jpeg")) {
    install.packages("jpeg")              
    library("jpeg")
}
if (!require("ggridges")) {
    install.packages("ggridges")              
    library("ggridges")
}
if (!require("plyr")) {
    install.packages("plyr")              
    library("plyr")
}
if (!require("ggiraph")) {
    install.packages("ggiraph")              
    library("ggiraph")
}
if (!require("highcharter")) {
    install.packages("highcharter")              
    library("highcharter")
}
if (!require("forecast")) {
    install.packages("forecast")              
    library("forecast")
}
## 
knitr::opts_chunk$set(echo = TRUE,       
                      warning = FALSE,   
                      result = TRUE,   
                      message = FALSE,
                      comment = NA)
```

# Data Summary
## Income/Life Data
  We first read in two data sets called "income" and "life" representing income and life expectancy values throughout a multitude of years. "Income" has 193 observations with 220 total variables whilst "Life" has 187 observations and 220 total variables.
```{r, comment=NA}
income <- read.csv("https://ecoleman451.github.io/ecoleman/w6/income_per_person.csv")

life <- read.csv("https://ecoleman451.github.io/ecoleman/w6/life_expectancy_years.csv")
```

## Reshape Income/Life Data
  Next, we reshape both data sets such that there are only three columns (Geo, Year, Income or Life Expectancy):
```{r, comment=NA}
# Reshape data set such that there are only three columns (Geo, Year, & Income)
new_income <- pivot_longer(income, cols = -geo, names_to = "year", values_to = "income")
head(new_income)

new_life <- pivot_longer(life, cols = -geo, names_to = "year", values_to = "life.expectancy")
head(new_life)
```

## Create New Data Set
  We then merge these two new sets into a data set called "LifeExpIncom" which now contains Geo, Year, Income, & Life Expectancy (40953 observations and 4 variables).
```{r, comment=NA}
## Create new data set
LifeExpIncom <- merge(new_life, new_income, by = c("geo", "year"))
head(LifeExpIncom)
```

## Adding in Country/Population Data
  We then read in two more sets called "country" (240 observations and 11 variables) and "pop" (195 observations and 220 variables) respectively representing country and population data. We reshape the data set "pop" so that it coincides with "LifeExpIncom" and "Country" which already have the variable Year transformed into one column.
```{r, comment=NA}
## Read in More Data
country <- read.csv("https://ecoleman451.github.io/ecoleman/w6/countries_total.csv")

pop <- read.csv("https://ecoleman451.github.io/ecoleman/w6/population_total.csv")

new_pop <- pivot_longer(pop, cols = -geo, names_to = "year", values_to = "population")
head(new_pop)

## Merge LifeExpIncom with Country
merged <- merge(LifeExpIncom, country, by.x = "geo", by.y = "name", all.x = TRUE)
head(merged)
```

## Creating Final Data Set
  After doing this, we're able to merge "LifeExpIncom" with "Country" and then this newly merged set with our recently transformed "pop" set, creating a set called "fin_data" (42705 observations and 15 variables).
```{r, comment=NA}
## Merge Population with Merged Data
fin_data <- merge(new_pop, merged, by = c("geo", "year"), all.x = TRUE)
head(fin_data)
```

## Subsetting Data for Year 2015
  After this, all that is left is to subset the data so that we only focus on data from the year 2015. This gives us our "final_data" (195 observations and 15 variables) set. Now, let's look at the overall summary statistics for the data set "fin_data" which contains not just data from 2015, but from all years from the data set.
```{r, comment=NA}
## Get Data for Year 2015
final_data <- subset(fin_data, year =="X2015")
summary(fin_data)
```

# Plotly
```{r, comment=NA}
scatter_plot <- plot_ly(
  data = final_data,
  x = ~income,
  y = ~life.expectancy,
  size = ~population,
  color = ~geo,
  text = ~paste("Country: ", geo, "<br>Population: ", population),
  type = "scatter",
  mode = "markers",
  marker = list(
    opacity = 0.6,  ## Transparency level
    sizemode = "diameter",  ## Set the size mode to diameter
    sizeref = 0.1,  ## Adjust the size reference for better visibility
    line = list(
      color = "black",  ## Boundary color for points
      width = 1  ## Boundary width
    )
  )
)
layout <- list(
  title = "Association Between Life Expectancy and Income (Year 2015)",
  xaxis = list(title = "Income"),
  yaxis = list(title = "Life Expectancy"),
  showlegend = FALSE  ## Hide legend for individual countries
)

## Combine the plot and layout
scatter_plot <- layout(scatter_plot, layout)

## Display the interactive scatter plot
scatter_plot
```

The above plot shows the relationship between income, life expectancy, and population size across different countries in the year 2015. Each point is a country and the size of the points correlate with the population size of that specific country. The countries are each color coded as well.

The x-axis looks at the income levels for each country. Countries that have higher incomes will be skewed to the right. The y-axis looks at life expectancy. Countries with higher life expectancy will be skewed higher on the y axis. From looking at the plot, we can see that there are some countries that primarily take over the scatter plot as opposed to others depending on population size and income. We can look at whether countries that have higher incomes generally have longer life expectancies or examine the population sizes to see if they correlate with higher or lower income levels.
