---
title: "Directory of Visualization"
author: "Edward Coleman"
date: "02-22-2024"
output:
  html_document: 
    toc: yes
    toc_depth: 4
    toc_float: yes
    fig_width: 6
    number_sections: yes
    theme: readable
    fig_height: 4
  pdf_document: 
    toc: yes
    toc_depth: 4
    fig_caption: yes
    number_sections: yes
    fig_width: 5
    fig_height: 4
---
<style type="text/css">

div#TOC li {
    list-style:none;
    background-color:lightgray;
    background-image:none;
    background-repeat:none;
    background-position:0;
    font-family: Arial, Helvetica, sans-serif;
    color: #780c0c;
}

/* mouse over link */
div#TOC a:hover {
  color: red;
}

/* unvisited link */
div#TOC a:link {
  color: blue;
}



h1.title {
  font-size: 24px;
  color: Darkblue;
  text-align: center;
  font-family: Arial, Helvetica, sans-serif;
  font-variant-caps: normal;
}
h4.author { 
    font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkRed;
  text-align: center;
}
h4.date { 
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
  text-align: center;
}

h1 {
    font-size: 22px;
    font-family: "Times New Roman", Times, serif;
    color: darkred;
    text-align: center;
}
h2 { 
    font-size: 18px;
    font-family: "Times New Roman", Times, serif;
    color: navy;
    text-align: left;
}

h3 { 
    font-size: 15px;
    font-family: "Times New Roman", Times, serif;
    color: navy;
    text-align: left;
}

h4 { 
    font-size: 18px;
    font-family: "Times New Roman", Times, serif;
    color: darkred;
    text-align: left;
}

/* unvisited link */
a:link {
  color: green;
}

/* visited link */
a:visited {
  color: green;
}

/* mouse over link */
a:hover {
  color: red;
}

/* selected link */
a:active {
  color: yellow;
}

</style>


```{r setup, echo = FALSE, message = FALSE, include = FALSE, warning=FALSE}
# run setup script
#install.packages("remotes")
#library(remotes)
#remotes::install_github("clauswilke/dviz.supp")
#devtools::install_github("clauswilke/dviz.supp")

library(colorspace)
library(dplyr)
library(tidyverse)
library(ggforce)
library(ggridges)
library(treemapify)
library(forcats)
library(statebins)
library(sf)
library(cowplot)


options(digits = 3)
knitr::opts_chunk$set(
                 echo = FALSE,
              message = FALSE,
              warning = FALSE,
               cache = FALSE,
               #dpi = 105, # not sure why, but need to divide this by 2 to get 210 at 6in, 
                           # which is 300 at 4.2in
           fig.align = 'center',
           fig.width = 6,
             fig.asp = 0.618,  # 1 / phi
            fig.show = "hold"
            )
options(dplyr.print_min = 6, dplyr.print_max = 6)
```

# Reading in Data Sets
```{r}
income <- read.csv("https://ecoleman451.github.io/ecoleman/w5/income_per_person.csv")

life <- read.csv("https://ecoleman451.github.io/ecoleman/w5/life_expectancy_years.csv")
```
```{r}
# Reshape data set such that there are only three columns (Geo, Year, & Income)
new_income <- pivot_longer(income, cols = -geo, names_to = "year", values_to = "income")
head(new_income)

new_life <- pivot_longer(life, cols = -geo, names_to = "year", values_to = "life.expectancy")
head(new_life)
```

## Create new data set
```{r}
LifeExpIncom <- merge(new_life, new_income, by = c("geo", "year"))
head(LifeExpIncom)
```

## Read in More Data
```{r}
country <- read.csv("https://ecoleman451.github.io/ecoleman/w5/countries_total.csv")
head(country)

pop <- read.csv("https://ecoleman451.github.io/ecoleman/w5/population_total.csv")

new_pop <- pivot_longer(pop, cols = -geo, names_to = "year", values_to = "population")
head(new_pop)
```

## Merge LifeExpIncom with Country
```{r}
merged <- merge(LifeExpIncom, country, by.x = "geo", by.y = "name", all.x = TRUE)
head(merged)
```

## Merge Population with Merged Data
```{r}
final_data <- merge(new_pop, merged, by = c("geo", "year"), all.x = TRUE)
head(final_data)
```

# GGPlots
## Scatter Plots
```{r}
scatter_pop <- ggplot(final_data, aes(x = year, y = population, color = region, size = life.expectancy)) +
  geom_point() +
  labs(title = "Yearly Population Data",
       x = "Year",
       y = "Population",
       size = "Life Expectancy",
       color = "Region")

scatter_life <- ggplot(final_data, aes(x = year, y = income, color = region, size = life.expectancy)) +
  geom_point() +
  labs(title = "Yearly Income Data",
       x = "Year",
       y = "income",
       size = "Life Expectancy",
       color = "Region")

combined_plot <- plot_grid(scatter_pop, scatter_life)
combined_plot
```

## Another GGPlot
```{r}
year = final_data$year
# Create a ggplot
ggplot(final_data, aes(x = year)) +
  geom_line(aes(y = population, color = "Population"), size = 1) +
  geom_line(aes(y = life.expectancy*1e6, color = "Life Expectancy"), size = 1) + # Multiplying life expectancy by 1e6 for better scale
  scale_y_continuous(sec.axis = sec_axis(~./1e6, name = "Life Expectancy (millions)")) +
  labs(title = "Population and Life Expectancy Over Years",
       x = "Year",
       y = "Population",
       color = "Variable") +
  theme_minimal()
```













