---
title: "Penguins 2"
output:
  html_document:
    code_folding: hide
    highlight: textmate
date: "2024-02-15"
---

<h1 class="title toc-ignore">Getting Started with Base R Graphics</h1>
<h4 class="author">Edward Coleman</h4>
<h4 class="date">West Chester University</h4>

</div>


<style type="text/css">

div#TOC li {
list-style:none;
background-color:lightgray;
background-image:none;
background-repeat:none;
background-position:0;
font-family: Arial, Helvetica, sans-serif;
color: #780c0c;
}

h1.title {
font-size: 24px;
color: darkblue;
text-align: center;
font-family: Arial, Helvetica, sans-serif;
font-variant-caps: normal;
}
h4.author { font-size: 18px;
font-family: Arial, Helvetica, sans-serif;
color: navy;
text-align: center;
}
h4.date { font-size: 18px;
font-family: Arial, Helvetica, sans-serif;
color: darkblue;
text-align: center;
}

h1 {
font-size: 22px;
font-family: "Times New Roman", Times, serif;
color: darkred;
text-align: left;
}
h2 {
font-size: 18px;
font-family: "Times New Roman", Times, serif;
color: navy;
text-align: left;
}
h3 { font-size: 15px;
font-family: "Times New Roman", Times, serif;
color: darkred;
text-align: left;
}
h4 {
font-size: 18px;
font-family: "Times New Roman", Times, serif;
color: darkred;
text-align: left;
}


a:link {
color: green;
}

a:visited {
color: purple;
}

a:hover {
color: red;
}

a:active {
color: yellow;
}
</style>
<div id="base-plotting-functions" class="section level1" number="1">
<h1><span class="header-section-number">1</span> First Plots</h1>
<div class="figure" style="text-align: center">
```{r}
pen <- read.csv("w03-penguins.csv")
# Delete all records with at least one missing componenet
new_pen <- na.omit(pen)

# Include Adelie & Gentoo Penguins from Biscoe & Torgersen Islands
species = new_pen$species
islands = new_pen$island

adelie.id = which(species =="Adelie")
gentoo.id = which(species =="Gentoo")
biscoe.id = which(islands =="Biscoe")
torg.id = which(species =="Torgersen")

# Include only penguins with body_mass_g less than 5000 and more than 3500
penguin <- pen[c(biscoe.id, torg.id, adelie.id, torg.id),]
penguin <- penguin[penguin$body_mass_g > 3500 & penguin$body_mass_g < 5000,]

# Rescale body_mass_g to BMI (body_mass_g/4000)
penguin$BMI <- penguin$body_mass_g/4000

# Exclude variables (X, sex, year, and body_mass_g)
new <- subset(penguin, select = -c(X, sex, year, body_mass_g))

plot(new$bill_length_mm, new$flipper_length_mm)
title(main = "Penguin Bill & Flipper Lengths")
```

<div class="figure" style="text-align: center">
<p class="caption">
Base plot with annotation: Iris
</p>
</div>
<p>Here, I start with the same plot as above (although I add the title
right away using the main argument to <code>plot()</code>) and then
annotate it by coloring blue the data points corresponding.</p>
```{r}
bill_length = new$bill_length_mm
flipper_length = new$flipper_length_mm
species = new$species

adelie.id = which(species =="Adelie")
gentoo.id = which(species =="Gentoo")

plot(bill_length, flipper_length, main = "Penguin Bill & Flipper Length")
points(bill_length[adelie.id], flipper_length[adelie.id], pch=19, col="red")
points(bill_length[gentoo.id], flipper_length[gentoo.id], pch=19, col="green")
```

</div>
<div class="figure" style="text-align: center">
<div id="controlling-point-size-and-transparency" class="section level1" number="3">
<h1><span class="header-section-number">3</span> Final Plot (Controlling Point Size
and Transparency)</h1>
<p>we use <code>cex=</code> and <code>alpha=</code> to control the point
size according to the value of a variable and the level of transparency
of the point.</p>
```{r}
bill_length = new$bill_length_mm
flipper_length = new$flipper_length_mm
species = new$species
body_mass = (new$body_mass_g/5000)

adelie.id = which(species =="Adelie")
gentoo.id = which(species =="Gentoo")

plot(bill_length, flipper_length, main = "Bill Length vs Flipper Length by Body Mass")
points(bill_length[adelie.id], flipper_length[adelie.id], 
       pch = 19, col = "red", cex = body_mass[adelie.id])
points(bill_length[gentoo.id], flipper_length[gentoo.id], 
       pch = 19, col = "green", cex = body_mass[gentoo.id])
legend("topleft", c("adelie", "gentoo"), col = c("red", "green"), pch=c(18, 19, 20))
```